getdata3.Rmd

### 2022-02-02 getdata2.Rmd
updates

```{r gettemp}
source("R/tempdat.R")
temp_start <- 1894
temp_end <- 1951
temp_ivl <- c(temp_start, temp_end)
##
umetemp <- tempdat("ume", temp_ivl)
bjurtemp <- tempdat("bjur", temp_ivl)
##stentemp <- tempdat("~/Forskning/Data/sten_temp.csv", c(1899, 1952))
saveRDS(umetemp, file = "data/umetemp.rds")
saveRDS(bjurtemp, file = "data/bjurtemp.rds")
```

>>>>

```{r gettemp}
source("R/tempdat.R")
temp_start <- 1894
temp_end <- 1951
temp_ivl <- c(temp_start, temp_end)
##
umetemp <- tempdat("ume", temp_ivl)
bjurtemp <- tempdat("bjur", temp_ivl)
stentemp <- tempdat("sten", temp_ivl) # Added 27 Jan 2022
##stentemp <- tempdat("~/Forskning/Data/sten_temp.csv", c(1899, 1952))
saveRDS(umetemp, file = "data/umetemp.rds")
saveRDS(bjurtemp, file = "data/bjurtemp.rds")
saveRDS(stentemp, file = "data/stentemp.rds")  ## Added 27 Jan2022
```

### 2022-02-02 getdata3.Rmd
use lag=0 (no lag)
updates

Now (December 22, 2021) I have come to the conclusion that data from the data frame 
*persons* is the best and easiest start.

So we start with *skum::persons* instead of *skum::obs*, and age is to begin with in days,
with stillbirths age *removed* and born alive but dead on birthday at age 0.25 days.

@@

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
per <- per[(per$subreg != "sten"), ]
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
saveRDS(per, file = "data/infdat3.rds")
```

@@

```{r makecommunal}
per$enter <- per$enter / 365
per$exit <- per$exit / 365
## per$birthdate <- toTime(per$birthdate) #already done!
temp_start <- 1894
temp_end <- 1951
temp_ivl <- c(temp_start, temp_end)
##
umetemp <- readRDS("data/umetemp.rds")
bjurtemp <- readRDS("data/bjurtemp.rds")
##

lagg <- 3 / 365  ## NOTE!!
tempvars <-  c("heat", "aver", "cold", "emintemp", "emeantemp", "emaxtemp", 
               "week", "year") 
## Ume:
bume <- per[per$subreg == "ume", ]
bume <- make.communal(bume, umetemp[, tempvars], 
                      start = temp_start, period = 1/52, lag = lagg)

@@

bume$region <- "ume"
bskel$region <- "ske"
infdat <- rbind(bume, bskel)
infdat <- infdat[order(infdat$id, infdat$enter), c("region", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socStatus", "hisclass", "illeg",
                                                   "parity", tempvars)]
saveRDS(infdat, file = "data/infdat3_year3.rds")

@@

>>>>

**Revisions:**

1.    (December 22, 2021) I have come to the conclusion that data from the data frame 
*persons* is the best and easiest start.

2.    (January 27. 2022) I found that teperature reports from *Stensele* are very similar
to those from the very similar *Umeå* and *Skellefteå*. Conclusion: The Skellefteå inland is now included.

We start with *skum::persons* instead of *skum::obs*, and age is to begin with in days,
with stillbirths age *removed* and born alive but dead on birthday at age 0.25 days.

@@

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "unknown"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "unknown"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "enter", "exit", "event")]
#saveRDS(per, file = "data/infdat3.rds")
```

@@

```{r makecommunal}
per$enter <- per$enter / 365
per$exit <- per$exit / 365
## per$birthdate <- toTime(per$birthdate) #already done!
temp_start <- 1894
temp_end <- 1951
temp_ivl <- c(temp_start, temp_end)
##
umetemp <- readRDS("data/umetemp.rds")
bjurtemp <- readRDS("data/bjurtemp.rds")
stentemp <- readRDS("data/stentemp.rds")
##

lagg <- 0 / 365 ##3 / 365  ## NOTE!!
tempvars <-  c("heat", "heat.1", "aver", "aver.1", "cold", "cold.1", "emintemp", "emeantemp", "emaxtemp", 
               "week", "year") 
## Ume:
bume <- per[per$subreg == "ume", ]
bume <- make.communal(bume, umetemp[, tempvars], 
                      start = temp_start, period = 1/52, lag = lagg)

@@

### Sten:
bsten <- per[per$subreg == "sten", ]
bsten <- make.communal(bsten, stentemp[, tempvars], 
                       start = temp_start, period = 1/52, lag = lagg)

bume$region <- "ume"
bskel$region <- "ske"
bsten$region <- "sten"
infdat <- rbind(bume, bskel, bsten)
infdat <- infdat[order(infdat$id, infdat$enter), c("subreg", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socBranch", "socStatus", "illeg",
                                                   "parity", "hisclass", tempvars)]
## New addition: (28 Jan 2022)
infdat <- rc(infdat) # rc from 'skum'
n <- NROW(infdat)

saveRDS(infdat, file = "data/infdat3.rds") # NOTE: Name change!!!

@@

Note the name change of the data file: *infdat3.rds*! And the the time lag is `r lagg`.
